### 2025-01-20 Thesis/Latex/Abstract.tex
Organizing repo and adding thesis template

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[arabic, english, french]{babel}
\begin{document}
	\newpage
	\thispagestyle{empty}
	\begin{center} \end{center}
	\vspace{0.01cm}

	\selectlanguage{arabic}
	\begin{center} {\LARGE ملخص  } \end{center}
	\par ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص ملخص .\\
	\textbf{\textit{كلمات مفتاحية:}} \textit{ملخص, ملخص, ملخص, ملخص, ملخص.}

	\vspace{2cm}

	\selectlanguage{english}
	\begin{center} {\LARGE Abstract} \end{center}
	\par Abstract Abstract Abstract  Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract.\\
	\textbf{\textit{Key words:}} \textit{Abstract, Abstract, Abstract, Abstract, Abstract.}

	\thispagestyle{empty}
\end{document}