### 2025-01-20 Thesis/Latex/Chapter_6.tex

\chapter{Seq2Seq Neural Machine Translation}
\pagestyle{fancy}\lhead{\textbf \footnotesize\it{Seq2Seq Neural Machine Translation}}
\pagestyle{fancy}\chead{} \pagestyle{fancy}\rhead{}
\pagestyle{fancy}\lfoot{\textbf {\small\it{Univ-Mascara/Computer Science: 2024}}} 
\pagestyle{fancy}\cfoot{} \pagestyle{fancy}\rfoot{\thepage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}\label{start6}

\section{System Architecture}
\section{Baseline System}
\section{Experiments and Evaluation}
\section{Results and Discussions}

\section{Summary}