### 2025-01-20 Thesis/Latex/Dedication.tex

\thispagestyle{empty}
\begin{tikzpicture}[remember picture, overlay]
	\node at (current page){
		\includegraphics[width=\paperwidth,height=\paperheight]{Figures/bgg4.png}} ;
\end{tikzpicture}
\vskip15mm
\begin{center}
	\resizebox{6cm}{1.1cm}{\calligra Dedication} 
\end{center}
%

\thispagestyle{empty}