### 2025-01-20 Thesis/Latex/Images_sources.tex

\begin{landscape}
	\thispagestyle{empty}
	\begin{center}
		\Huge{\textbf{Sources of Images}} \\
		\normalsize{\textit{for Copyright purposes}}
	\end{center}
	Figure 2.1 from \url{www.site1.com} \\
	Figure 2.3 from \url{www.site2.fr} \\
	Figure 2.5 from \url{www.site3.com} \\
	Figure 2.9 from \url{www.site4.com} \\
	Figure 3.5 from \url{www.site5.com} \\
	Figure 2.3 from \url{www.site2.fr} \\
	Figure 2.5 from \url{www.site3.com} \\
	Figure 2.9 from \url{www.site4.com} \\
	Figure 3.5 from \url{www.site5.com} \\
\end{landscape}