### 2025-01-22 Thesis/Latex/Cover_page.tex

\thispagestyle{empty}
\begin{center}
	\begin{title}
		\textbf{\uppercase{ People's Democratic Republic of Algeria }} \\[0.2cm]
		\textbf{\uppercase{ MINISTRY OF HIGHER EDUCATION AND SCIENTIFIC RESEARCH }} \\[0.5cm]
		\textbf{\uppercase{ UNIVERSITY MUSTAPHA STAMBOULI OF MASCARA}}
	\end{title}
	\begin{figure}[H]
		\centering
		\label{logo_UD}\includegraphics[width=0.2\textwidth]{Figures/Logo_UM.jpg}
	\end{figure}
	{\large Faculty of Exact Sciences} \\[0.2cm]
	{\large Department of Computer Science} \\[0.2cm]
	{\large Dissertation } \\[0.2cm]
	{\ Submitted in partial fulfilment of the requirements for Master degree in Computer Science} \\[0.2cm]
	{\large Option: Artificial Intelligence}\\[0.5cm]

    {\large Theme} \\[0.3cm]
%\hline	
%    \\[0.7cm]
	{\LARGE {\textbf{Enhancing Legal Information Access and \\ Reasoning  with  Retrieval-Augmented LLMs for Juridical Data}}}
	\\[0.7cm]
%\hline
%	 \\[0.5cm]

	{\large Presented by \textbf{Boudjenane Zoubida Asmaa} }\\[0.5cm]

\end{center}
Jury:\\
\begin{tabular}{llll}
	President    &  Aaaaa AAAAA & Professor & University of Mascara\\
	Director     &  Mohammed SALEM & Professor & University of Mascara\\
%	Co-Director  &  Ccccc CCCCC & Professor & University of Mascara\\
	Examiner     &  Ddddd DDDDD & Professor	& University of Mascara\\
	Examiner	 &  Eeeee EEEEE & Associate Prof.  & University of Mascara\\
	Examiner	 &  Fffff FFFFF & Associate Prof.  & University of Mascara\\
\end{tabular}
\vfill
\begin{center}

	\it{Month  2025}
\end{center}