### 2025-04-15 Thesis/Latex/Acknowledgements.tex

\chapter*{Acknowledgements}
\thispagestyle{empty}
%\addcontentsline{toc}{chapter}{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}

	First of all, praise and thanks to Allah Almighty for giving us all the patience, courage, will and motivation that allowed us to accomplish this work.\\ 
	\vspace{0.3cm}

	My deepest gratitude goes to my supervisor, \textbf{Mr Salem Mohamed } for his patient guidance, insightful feedback, and continuous encouragement throughout the course of this work .\\
	\vspace{0.3cm}
	I would like to acknowledge the members of the jury for their interest  \\
	\vspace{0.3cm}
    I am also especially thankful to my parents, family, and friends for their unwavering support, love, and motivation at every step of this journey. \\
    \vspace{0.3cm}
    I would like to thank everyone who helps me to improve my work. and who gave me any remark that helped me to perfect this manuscript.\\
	\vspace{0.3cm}
	All the teachers of the University MUSTAPHA STAMBOULI DE LA WILAYA MASCARA For the good contribution of this work.
\end{center}