### 2025-04-15 Thesis/Latex/Dedication.tex

\thispagestyle{empty}
\begin{tikzpicture}[remember picture, overlay]
	\node at (current page){
		\includegraphics[width=\paperwidth,height=\paperheight]{Figures/bgg4.png}} ;
\end{tikzpicture}
\vskip15mm
\begin{center}
	\resizebox{6cm}{1.1cm}{\calligra Dedication} 
\end{center}
\vfill 
\begin{center}
	\parbox{.85\linewidth}{\centering 
		\baselineskip=8mm 
		\fontfamily{pzc}\selectfont 
		\Large 
This work is dedicated with deep love and gratitude to:
\\[5mm]
My dear parents,
whose unwavering support, sacrifices, and constant prayers have been the foundation of everything I’ve achieved. Your belief in me gave me the strength to persevere through every challenge.
\\[5mm]
My supervisor, Mr. [Mr Salem Mohamed],
whose patient guidance, insightful feedback, and continuous encouragement made this journey not only possible but meaningful. Your dedication, expertise, and belief in my potential have left a lasting impact on me both academically and personally.
\\[5mm]
My beloved sister,
for being my greatest cheerleader and my constant source of emotional support. Your encouragement and care during the toughest moments meant the world to me.
\\[5mm]
My wonderful friend [nial hadjer],
for always being there with kind words, late-night conversations, and unshakable support.
\\[19mm]
}
\end{center}
\vfill 
\thispagestyle{empty}